---
title: "206_Assignment_5"
author: "Tess Hooper"
date: "12/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####################################
####################################

## Step 1. Load packages and read in data.
```{r, echo=FALSE, message=FALSE}

library(tidyverse)
library(stargazer)
library(knitr)
library(pwr)
library(plotly)
library(ggrepel)
library(effsize)
library(vcdExtra)
library(dplyr)
library(car)

grad_enroll <- read_csv("grad_enroll.csv")
median_salary <- read_csv("median_salary_doc.csv")
phd_field <- read_csv("phd_field.csv")
faculty <- read_csv("faculty_salary.csv")

```



####################################
####################################

Part 1: Male and female graduate enrollment (1967 - 2015)

## Compare trends in total graduate enrollment for males and females (including full-time/part-time and private/public universities) in the United States from 1967 - 2015. Describe your results statistically, graphically and in text.

a. Exploratory Data Analysis
- Summary data table with graduate enrollment for males and females, part-time/full-time, private/public universites
- Scatter plots for full-time/part-time and gender, private/public universities and gender 
- Regression line for males and females separately, then put together on the same graph (we need to do a regression because we need the line of best fit to find residuals in order to make sure our model fits our assumptions - residuals are any point that isn't on the line of best fit )

b. Question: How does enrollment in regards to gender change within these categories over time?


```{r, echo=FALSE, message = FALSE}

# Exploratory scatter plot for total male and female grad enrollment by year

grad_scatter <- ggplot(grad_enroll, aes(x = Year, y = Total_males)) +
  geom_point()+
  geom_point(aes(x = Year, y = Total_females))

grad_scatter

# Exploratory summary table 

grad_table <- 

#geom_smooth (method = lm, se = F, aes (col = "Male Enrollment"))

```


Part 2. Shifts in female PhD recipients by field (1985, 2000, and 2015). 

a. Describe if and how there was a shift in PhDs awarded to females in Physical and Earth Sciences, Engineering, Education,
and Humanities & Arts in 1985, 2000, and 2015. 

b. Describe in a graph or table.

c. Describe in text. 

#Note: several ways to interpret this question: decide which you think is/are most interesting. state which you are asking/answering in your report.




##Part 3: Male and female salaries for starting postdoctoral and other employment positions (2015).

Compare median salaries for male and female doctorate recipients in 2015. Answer these two questions:

A. Does median salary differ significantly between male and female starting postdoc positions?

B. Does median salary differ significantly between male and female PhD recipients in non-postdoc employment positions?




