---
title: "206_Assignment_5"
author: "Tess Hooper"
date: "12/2/2018"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####################################
####################################

## Step 1. Load packages and read in data.
```{r, echo=FALSE, message=FALSE}

library(tidyverse)
library(stargazer)
library(knitr)
library(pwr)
library(plotly)
library(ggrepel)
library(effsize)
library(vcdExtra)
library(dplyr)
library(car)
library(stargazer)

grad_enroll <- read_csv("grad_enroll.csv")
median_salary <- read_csv("median_salary_doc.csv")
phd_field <- read_csv("phD_field.csv")
faculty <- read_csv("faculty_salary.csv")

```



####################################
####################################

Part 1: Male and female graduate enrollment (1967 - 2015)

## Compare trends in total graduate enrollment for males and females (use Totals) in the United States from 1967 - 2015. Describe your results statistically, graphically and in text.

a. Exploratory Data Analysis
- Summary data table with graduate enrollment for males and females, part-time/full-time, private/public universites
- Scatter plots for full-time/part-time and gender, private/public universities and gender 
- Regression line for males and females (we need to do a regression because we need the line of best fit to find residuals in order to make sure our model fits our assumptions - residuals are any point that isn't on the line of best fit )

b. Question: How does enrollment in regards to gender change within these categories over time?


```{r, echo=FALSE, message = FALSE}

### Read in grad_enrollment csv

grad_enroll <- read_csv("grad_enrollment.csv")

### Exploratory scatter plot for total male and female grad enrollment by year

grad_scatter <- ggplot(grad_enroll, aes(x = Year, y = Total_Males, col = "Males")) +
  geom_point()+
  geom_point(aes(x = Year, y = Total_Females, col = "Females"))

grad_scatter # The data looks like a linear model makes sense


### Conduct linear model 

female_lm <- lm(grad_enroll$Total_Females ~ grad_enroll$Year)
female_lm
# intercept = -5895502 (y-intercept)
# year = 30126 (slope)

male_lm <- lm(grad_enroll$Total_Males ~ grad_enroll$Year)
male_lm
# intercept = -17112153 (y-intercept)
# year = 9069 (slope)
# This model shows us that for both males and females, enrollment is increasing each year

plot(male_lm) # data is normally distributed
plot(female_lm) # data is normally distributed 


### Explore model fit and significance

summary(female_lm) 
# b = -5895502
# m = 30126

# y = 30126x - 5895502 (y = grad enrollment)
# Correlation coefficeint = 0.9827 (adjusted R-squared = 0.9823)
# p-value = < 2.2e-16
## Our R-squared value shows that there is strong model fit (meaning that using year against enrollment is a good idea)

summary(male_lm)
# b = -17112153
# m = 9069

# y = 9069x - 17112153 (y = grad enrollment)
# Correlation coefficient = 0.8545 (adjusted R-squared = 0.8514)
# p-value = < 2.2e-16
## Our R-squared value shows that there is strong model fit (meaning that using year against enrollment is a good idea)



### Plot data with regression line on graph -- FINALIZE GRAPH FOR REPORT

grad_graph <- ggplot(grad_enroll, aes(x = Year, y = Total_Males, col = "Males")) +
  geom_smooth(method = lm, se = F, aes(col = "Male")) +
  geom_smooth(aes(x = Year, y = Total_Females, col = "Female"), method = lm, se = F) +
  ylab("Enrollment") +
  ggtitle("Changes in Graduate University Enrollment by Sex (1967-2015)")+
  scale_color_discrete(name = "Gender")+
  theme_classic ()

grad_graph

### Find correlation using Pearson's R

female_pr <- cor.test(grad_enroll$Year, grad_enroll$Total_Females)
female_pr

# Pearson's R = 0.9913086
# There is a strong/positive correlation between year and total female enrollment

male_pr <- cor.test(grad_enroll$Year, grad_enroll$Total_Males)
male_pr

# Pearson's R = 0.9243741
# There is a strong/positive correlation between year and total male enrollment


### Explain your results in words:

# FEMALES: Year significantly predicts total female enrollment in graduate level university (b = -5895502, t(47) = 2669, p < 2.2e-16) with a strong positive correlation between the two (Pearson’s r = 0.99). The overall model (total female enrollment = 30126*(year) - 5895501, in count of students) explains a significant amount of variance in total female graduate enrollment (F(1,47) = 2669, p < 2.2e-16, R2 = 0.98).

# FEMALES SIMPLE: The model (total female enrollment = 30126*(year) - 5895501, in count of students) explains a significant amount of variance in total female graduate enrollment (F(1,47) = 2669, p < 2.2e-16, R2 = 0.98, Pearson's r = 0.99). 


# MALES: Year significantly predicts total male enrollment in graduate level university (b = -17112153, t(47) = 276, p < 2.2e-16) with a strong positive correlation between the two (Pearson’s r = 0.92). The overall model (total male enrollment = 9069*(year) - 17112153, in count of students) explains a significant amount of variance in total male graduate enrollment (F(1,47) = 276, p < 2.2e-16, R2 = 0.8545).

# MALES SIMPLE: The model (total male enrollment = 9069*(year) - 17112153, in count of students) explains a significant amount of variance in total male graduate enrollment (F(1,47) = 276, p < 2.2e-16, R2 = 0.8545, Pearson's r = 0.92).

stargazer(female_lm, male_lm, type= "text", out="test.htm")



```


<<<<<<< HEAD









































































































































##Part 2. Shifts in female PhD recipients by field (1985, 2000, and 2015). 
=======
Part 2. Shifts in female PhD recipients by field (1985, 2000, and 2015). 
>>>>>>> b400d8b7001a71f3d0af765920a26a74c22bc5a0

a. Describe if and how there was a shift in PhDs awarded to females in Physical and Earth Sciences, Engineering, Education,
and Humanities & Arts in 1985, 2000, and 2015. 

b. Describe in a graph or table.

c. Describe in text and statistically. 

#Note: several ways to interpret this question: decide which you think is/are most interesting. state which you are asking/answering in your report.

```{r, echo=FALSE, message = FALSE}
# Get the 4 fields and only females.

subject_fphds <- phd_field  %>%
  filter(field == "physical and earth sciences"| field == "Engineering"| field == "Education"| field == "Humanities and arts") %>% 
  filter(sex == "female") %>% 
  mutate(field = case_when(
    field == "physical and earth sciences" ~ "Physical Sciences",
    field == "Engineering" ~ "Engineering",
    field == "Education" ~ "Education",
    field == "Humanities and arts" ~ "Arts & Humanities"
  ))

subject_fphds

peryear <- subject_fphds  
  

#Line graph for female phDs across the fields

females_phd_graph <- ggplot(subject_fphds, aes(x=year, y=number, group=field)) +
  geom_line(aes(color=field))+
  geom_point(aes(color=field))+
  labs(x = "Year", y="Frequency (Female-Awarded PhD Degrees)", title = "Number of PhD Degrees Awarded to Females by Field")+
  scale_x_continuous(breaks = c(1985, 2000, 2015))+
  theme_classic()

females_phd_graph

Year85 <- c(569, 198, 3491, 1392)
Year00 <- c(1022, 838, 4179, 2672)
Year15 <- c(1988, 2301, 3502, 2832)

#85=5650
#00=8711
#15=10,623

chi_females <- rbind(Year85, Year00, Year15)
colnames(chi_females) <- c("Pysical Sciences", "Engineering", "Education", "Arts & Humanities")
```




```{r}
#1. Use chi-squared to explore changes in PhD proportions between fields over time. Are we testing ito see if there is a 

#Null hypothesis (H0): There is no significant association between number of degrees awarded and field of study
#Alternative hypothesis (Ha): There is a significant association between number of degrees awarded and field of study.

x2_test <- chisq.test(chi_females)


x2_test

#interested in showing the changes in proportions of number of degrees between the categories across the 3 years. We also want to show total increase or decrease in number of PhD degrees per category.


###p-value is <0.001. There IS a significant association between number of degrees awarded and field of study. 
### x2(6)= 2073.3, p<0.001 

```





























































































































































































































   








##Part 3: Male and female salaries for starting postdoctoral and other employment positions (2015).

Compare median salaries for male and female doctorate recipients in 2015. Answer these two questions:

A. Does median salary differ significantly between male and female starting postdoc positions?

B. Does median salary differ significantly between male and female PhD recipients in non-postdoc employment positions?


```{r, echo=FALSE, message=FALSE}

#Ways to explore the medians: Mann-Whitney U and Wilcoxen Signed Rank
#Mann-Whiteney U is the non paired test

#Explore the Data with histograms

postdoc_hist_male <- ggplot(median_salary, aes(x= postdoc_male)) +
  geom_histogram()

postdoc_hist_male


postdoc_hist_female <- ggplot(median_salary, aes(x = postdoc_female))+
  geom_histogram()

postdoc_hist_female

postdoc_qq_male <- ggplot(median_salary, aes(sample = postdoc_male))+
  geom_qq()

postdoc_qq_male

postdoc_qq_female <- ggplot(median_salary, aes(sample = postdoc_female)) +
  geom_qq()

postdoc_qq_female


#no assupmtion of normality necessary, no outliers 


#Wilcox test because data is paired

#H0: The medians will not be signifigantly different between male and female postdoc salaries

#HA: The medians will be signifigantly different between male and female postdoc salaries 

postdoc_salary_wilcox <- wilcox.test(median_salary$postdoc_male, median_salary$postdoc_female, paired = TRUE)

postdoc_salary_wilcox

#V = 19.5, p-value = 0.8884
#There is no signifigant difference between males and females in postdoctural positions

#Employment in non-postdoc positions 
employment_salary_wilcox <- wilcox.test(median_salary$employment_male, median_salary$employment_female, paired = TRUE)

employment_salary_wilcox 

#V = 101, p-value = 0.002572
#There is a signifigant difference between males and females in non-postdoctural postions






```










##Part 4. Exploring academic salaries for professors in U.S. colleges.

Develop a model describing faculty salary based on data for faculty sex, rank, years in current position, field, and number of years since doctoral degree was earned.


```{r}

#exploratory graphs about prices by the other varaibles
#Salary based on sex
by_sex <- ggplot(faculty, aes(x = salary)) +
  geom_density(aes(color = sex, fill = sex), alpha = 0.3) + 
  theme_classic() +
  scale_x_continuous(expand = c(0,0), limits = c(0,300000)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "Professor Salary", y = "Density")

by_sex

#Salary based on years serviced 
by_years_service <- ggplot(faculty, aes(x = years_service, y = salary)) +
  geom_point(aes(color = sex), alpha = 0.5, scale = "free") +
  facet_wrap(~sex)

by_years_service

#Salary based on postgrad years

by_postgrad_years <- ggplot(faculty, aes(x = postgrad_years, y = salary)) +
  geom_point(aes(color = sex), alpha = 0.5, scale = "free") +
  facet_wrap(~sex)

by_postgrad_years


#saturated linear regression model from all variables

faculty_lrm1 <- lm(salary ~ rank + discipline + postgrad_years + years_service + sex, data = faculty)

faculty_lrm1

summary(faculty_lrm1)

#reference levels: Associate Professor, discipline A (applied) (B = Theoretical), and Female
#If everything else about the ranks are the same, I would expect an Assistant Professor to make $13000 less than an Associate Professor. And I would expect a full Professor to make $32000 more than an Associate Professor
#If everything else about the Discipline is the same, a faculty member in Applied would make $14500 more per year than a faculty member in Theoretical. 
#If everything else abou the sex is the same, I would expect a Male faculty member would make $4783 more than a Female faculty member. 
#For every year after postgrad increased, the salary would increase by $535
#The value for Service years doesn't make sense because this means: for every year working, the salary would decrease by $490

plot(faculty_lrm1)

#Going to make a model with out Service years to see if it makes sense. 

faculty_lrm2 <- lm(salary ~ rank + discipline + postgrad_years + sex, data = faculty)

faculty_lrm2

summary(faculty_lrm2)

#compare the models with AIC

lrm1_aic <- AIC(faculty_lrm1)
lrm2_aic <- AIC(faculty_lrm2)

#lrm1 is only 4 values less than lrm2, so we can stay with lrm1 with knowing that the service years is negative

#stargazer for the table to compare the models 
#open up test.htm to copy and edit in word
stargazer(faculty_lrm1, faculty_lrm2, type = 'html', out = "test.htm")

```

