---
title: "206_Assignment_5"
author: "Tess Hooper"
date: "12/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####################################
####################################

## Step 1. Load packages and read in data.
```{r, echo=FALSE, message=FALSE}

library(tidyverse)
library(stargazer)
library(knitr)
library(pwr)
library(plotly)
library(ggrepel)
library(effsize)
library(vcdExtra)
library(dplyr)
library(car)
library(stargazer)

grad_enroll <- read_csv("grad_enroll.csv")
median_salary <- read_csv("median_salary_doc.csv")
phd_field <- read_csv("phd_field.csv")
faculty <- read_csv("faculty_salary.csv")

```



####################################
####################################

Part 1: Male and female graduate enrollment (1967 - 2015)

## Compare trends in total graduate enrollment for males and females (use Totals) in the United States from 1967 - 2015. Describe your results statistically, graphically and in text.

a. Exploratory Data Analysis
- Summary data table with graduate enrollment for males and females, part-time/full-time, private/public universites
- Scatter plots for full-time/part-time and gender, private/public universities and gender 
- Regression line for males and females (we need to do a regression because we need the line of best fit to find residuals in order to make sure our model fits our assumptions - residuals are any point that isn't on the line of best fit )

b. Question: How does enrollment in regards to gender change within these categories over time?


```{r, echo=FALSE, message = FALSE}

### Read in grad_enrollment csv

grad_enroll <- read_csv("grad_enrollment.csv")

### Exploratory scatter plot for total male and female grad enrollment by year

grad_scatter <- ggplot(grad_enroll, aes(x = Year, y = Total_Males, col = "Males")) +
  geom_point()+
  geom_point(aes(x = Year, y = Total_Females, col = "Females"))

grad_scatter # The data looks like a linear model makes sense


### Conduct linear model 

female_lm <- lm(grad_enroll$Total_Females ~ grad_enroll$Year)
female_lm
# intercept = -5895502 (y-intercept)
# year = 30126 (slope)

male_lm <- lm(grad_enroll$Total_Males ~ grad_enroll$Year)
male_lm
# intercept = -17112153 (y-intercept)
# year = 9069 (slope)
# This model shows us that for both males and females, enrollment is increasing each year

plot(male_lm) # data is normally distributed
plot(female_lm) # data is normally distributed 


### Explore model fit and significance

summary(female_lm) 
# b = -5895502
# m = 30126

# y = 30126x - 5895502 (y = salmon imports)
# Correlation coefficeint = 0.9827 (adjusted R-squared = 0.9823)
# p-value = < 2.2e-16
## Our R-squared value shows that there is strong model fit (meaning that using year against enrollment is a good idea)

summary(male_lm)

# Plot data with regression line on graph

grad_graph <- ggplot(grad_enroll, aes(x = Year, y = Total_Males, col = "Males")) +
  geom_smooth(method = lm, se = F, aes(col = "Male")) +
  geom_smooth(aes(x = Year, y = Total_Females, col = "Female"), method = lm, se = F) +
  ylab("Enrollment") +
  ggtitle("Changes in Graduate University Enrollment by Sex (1967-2015)")+
  scale_color_discrete(name = "Gender")+
  theme_classic ()

grad_graph


```


<<<<<<< HEAD











































































































































##Part 2. Shifts in female PhD recipients by field (1985, 2000, and 2015). 
=======
Part 2. Shifts in female PhD recipients by field (1985, 2000, and 2015). 
>>>>>>> b400d8b7001a71f3d0af765920a26a74c22bc5a0

a. Describe if and how there was a shift in PhDs awarded to females in Physical and Earth Sciences, Engineering, Education,
and Humanities & Arts in 1985, 2000, and 2015. 

b. Describe in a graph or table.

c. Describe in text. 

#Note: several ways to interpret this question: decide which you think is/are most interesting. state which you are asking/answering in your report.

```{r, echo=FALSE, message = FALSE}

#Select and filter for just female PhDs in P+E Sciences, Engineering, Ed, and Humanities for 1985, 2000 + 2015

fem_phds <- phd_field  %>% 
  select(field_sex, `1985_N`)
 # select(1985_N, 1985_P, 2000_N, 2000_P, 2015_N, 2015_P)


```







































































































































































































































##Part 3: Male and female salaries for starting postdoctoral and other employment positions (2015).

Compare median salaries for male and female doctorate recipients in 2015. Answer these two questions:

A. Does median salary differ significantly between male and female starting postdoc positions?

B. Does median salary differ significantly between male and female PhD recipients in non-postdoc employment positions?


```{r, , echo=FALSE, message=FALSE}

#Ways to explore the medians: Mann-Whitney U and Wilcoxen Signed Rank
#Mann-Whiteney U is the non paired test

#Explore the Data with histograms










```


