---
title: "206_Assignment_5"
author: "Tess Hooper"
date: "12/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####################################
####################################

## Step 1. Load packages
```{r, echo=FALSE, message=FALSE}

library(tidyverse)
library(stargazer)
library(knitr)
library(pwr)
library(plotly)
library(ggrepel)
library(effsize)
library(vcdExtra)
library(dplyr)
library(car)

```



####################################
####################################

Part 1: Male and female graduate enrollment (1967 - 2015)

## Compare trends in total graduate enrollment for males and females (including full-time/part-time and private/public universities) in the United States from 1967 - 2015. Describe your results statistically, graphically and in text.

a. Exploratory Data Analysis
- Summary data table with graduate enrollment for males and females, part-time/full-time, private/public universites
- Histograms for full-time/part-time and gender, private/public universities and gender (and QQplots for each
- Regression line for males and females separately, then put together on the same graph (we need to do a regression because we need the line of best fit to find residuals in order to make sure our model fits our assumptions - residuals are any point that isn't on the line of best fit )

b. Question: How does enrollment in regards to gender change within these categories over time?


```{r, echo=FALSE, message = FALSE}

grad_trends <- grad_enroll


```








